\singlespacing{\textbf{Efficient pollution abatement in electricity markets with intermittent renewable energy}}

### 2019-04-26 documents/draft.tex
Update draft
- revise definition of intermittency/reliability
- update last paragraph of intro

\author[]{Saketh Aleti}

\affil[]{\small{}}

@@

To highlight this difference, we must first contrast the terms  \textit{intermittency} and \textit{reliability}. By intermittency, we mean uncontrollable variation in output over time due to physical constraints; by reliability, we mean variation in output that is wholly unpredictable\footnote{By unpredictible, we mean a notion similar to that of \textit{Knightian uncertainity} \citep{Knight}. } and may be the result of failures that cannot be modeled stochastically. Consider wind energy as an example. The energy provided by a wind turbine necessarily varies over time due to differences in wind speed and direction. Moreover, the relationship between the force created by the wind and the amount of wind power generated must necessarily be a fixed, functional relationship; that is, \textit{ceterius paribus}, two identical wind turbines under identical wind conditions should generate the same amount of energy. Consequently, the output of wind power can be forecasted up to an error term.\footnote{\citet{Foley2012} provide a review of the literature on the forecasting of wind power.} Because no forecast of wind (or solar) power can be perfect, some level of stochastic intermittency may be left in the error term. At the same time, the error term also captures unreliability; for instance, failures in the circuitry of wind turbines cannot simply be forecasted with a time series. So, overall, we may think of the energy output of a renewable technology as consisting of three components: a knowable/forecastable component related to intermittency, a stochastic component related to intermittency, and an unpredictible component related to reliability.

While \citet{HH} model the stochastic component of intermittency, we model the predictible component. More specifically, Helm and Mier model the capacity of renewable energy equal to a baseline capacity multiplied by a uniform random variable between 0 and 1. This is a valid approach given that their model consists of a single period of generation. On the other hand, we neglect stochastic intermittency for parsimony and develop a general, multi-period model where the output per unit of each technology follows a known function.  We then narrow this to a two period, two technology model where our first technology represents fossil fuel energy with constant output in both periods while our second technology's output differs between each period. Additionally, while their model consists of two sets of consumers, one set which receives fixed price contracts and another paying real time prices, we consider one representative consumer. Specifically, this representative consumer's utility is given by CES function where each good is electricity consumption differentiated by time. Interestingly, this treatment of intermittency produces some key results similar to that of Helm and Mier.

@@

In order to estimate this $\sigma$, we further modify the previous equation. 
Firstly, note that we cannot observe the demand shifter $\alpha_{t,i}$ directly, so we add a set of controls that may be responsible for shifts in demand. Our regression is then
\begin{align*}
\ln (Z_{t, i} / Z_{ s, i}) &= -\sigma \ln (P_{t,i} / P_{s,i}) +  \gamma_{t,i} A_{t,i} + \gamma_{s,i} A_{s,i} + u_i
\end{align*}
where $A$ represents set of controls for changes in demand while $u_i$ is a normal error term. Note that the control  $A_{t,i}$ replaces $\sigma \ln(\alpha_{t,i})$ and likewise for the period $s$ term; this substitution is valid because the $\ln(\alpha_{t,i}) \in \mathbb{R}$ and the $\sigma$ is simply absorbed into the estimated coefficient for $\gamma_{t,i}$. For the demand controls themselves, we choose to use heating (HDD) and cooling degree days (CDD) due to the aggregation of the data. That is, a more general control such as average temperature would not be able to directly capture intramonthly changes in demand, since variation in temperature would be lost when aggregated; on the other hand, CDDs and HDDs directly represent daily deviations in temperature even when totaled for each month. Additionally, demand for electricity may rise over time. Hence, we include, as a control, the difference in months between time $t$ and $s$; this is represented by $\Delta_{t,s}$. Finally, this panel requires us to consider fixed effects for each state, so we use a fixed effects panel regression. In total, the demand equation is:
\begin{align*}
\ln (Z_{ t, i} / Z_{ s, i}) &= -\sigma \ln (P_{t,i} / P_{s,i}) +  \gamma_{t,i} A_{t,i} + \gamma_{s,i} A_{s,i} + \eta \Delta_{t,s} + u_i \\
&= -\sigma \ln (P_{t,i} / P_{s,i}) +  \gamma_{t,i} \left( CDD_{t,i} + HDD{t,i} \right) \\
&\qquad + \gamma_{s,i} \left( CDD_{s,i} + HDD{s,i} \right)  + \eta \Delta_{t,s} + u_i
\end{align*}

@@

>>>>

\author[]{Saketh Aleti\thanks{I would like to thank Gal Hochman for helpful comments and the Sun Grant for funding this project. }}

\affil[]{\small{Rutgers University}}

@@

To highlight this difference, we must first contrast the terms  \textit{intermittency} and \textit{reliability}. By intermittency, we predictable changes in output related to physical constraints. For instance, while wind energy output varies over time, if  we know the angle and speed of wind, we may precisely derive the quantity of energy that a wind farm generates.\footnote{\citet{Foley2012} provide a review of the literature on the forecasting of wind power. Over time, forecasts have gotten much more accurate allowing electricity grids to manage wind power intermittency ahead of time.}  At the same time, these forecasts may not be perfect and, as a result, we see unexpected variation in the output of a technology. This unpredictability may be better captured by the notion of reliability; specifically,  we refer to definition provided by the US Department of Energy's \citep{ORNL} -- ``Power reliability can be defined as the degree to which the performance of the elements in a bulk system results in electricity being delivered to customers within accepted standards and in the amount desired." This is more so related to the stochastic or unpredictable variation in the output of a technology. Unlike intermittency, this cannot be planned around with 100\% certainty. A practical example may be a wind turbine's systems failing. While we may know the chances of this occurring, it's not possible to know when it will occur ahead of time; consequently, this may result in a temporary reduction in the quantity and quality of the electricity delivered --  a loss of reliability. In short, when \citeauthor{HH} define renewable output as equal to a baseline capacity multiplied by a uniform random variable, their model is closer to one of reliability rather than intermittency.

On the other hand, we neglect reliability for parsimony and develop a two-technology, two-period model consisting of fossil and renewable energy. Then, we model intermittency as defined above by allowing the output of renewable energy to differ between periods according to known function. At the same time, we consider a representative consumer whose utility is given by CES function where each good is electricity consumption differentiated by time. This captures consumers' preferences for electricity consumption throughout the day as well as their willingness to substitute consumption across time. We then fit the parameters of this CES utility function using electricity consumption and price data for each US state. At the same time, we set parameters of our technologies (their cost and output) to those of coal and solar. Overall, this treatment of intermittency produces some key results similar to that of Helm and Mier. That is, we also find a Pigovian tax to be a solution to externalities along with an S-shaped adoption curve for renewables. But, moreover, we find the elasticity of substitution between fossils and renewables to be highly non-linear; this directly challenges the CES production structure assumption made by standard models of fossil and renewable energy. Finally, we discuss how a combination of carbon taxes and research subsidies may be more optimal than carbon taxes alone in the presence of intermittency.

@@

In order to estimate this $\sigma$, we further modify the previous equation. 
Firstly, note that we cannot observe the demand shifter $\alpha_{t,i}$ directly, so we must replace the $\alpha$ terms with a set of controls that may be responsible for shifts in demand. So, still in general terms, our regression equation is now
\begin{align*}
\ln (Z_{t, i} / Z_{ s, i}) &= -\sigma \ln (P_{t,i} / P_{s,i}) +  \gamma_{t,i} A_{t,i} + \gamma_{s,i} A_{s,i} + u_i
\end{align*}
where $A$ represents set of controls for changes in demand while $u_i$ is a normal error term. Note that the control  $A_{t,i}$ replaces $\sigma \ln(\alpha_{t,i})$ and likewise for the period $s$ term; this substitution is valid because the $\ln(\alpha_{t,i}) \in \mathbb{R}$ and the $\sigma$ is simply absorbed into the estimated coefficient for $\gamma_{t,i}$. For the demand controls themselves, we choose to use heating (HDD) and cooling degree days (CDD) due to the aggregation of the data. That is, a more general control such as average temperature would not be able to directly capture intramonthly changes in demand, since variation in temperature would be lost when aggregated; on the other hand, CDDs and HDDs directly represent daily deviations in temperature even when totaled for each month. Additionally, demand for electricity may rise over time. Hence, we include, as a control, the difference in months between time $t$ and $s$; this is represented by $\Delta_{t,s}$. Finally, this panel requires us to consider fixed effects for each state, so we use a fixed effects panel regression. In total, the demand equation is:
\begin{align*}
\ln (Z_{ t, i} / Z_{ s, i}) &= -\sigma \ln (P_{t,i} / P_{s,i}) +  \gamma_{t,i} A_{t,i} + \gamma_{s,i} A_{s,i} + \eta \Delta_{t,s} + u_i \\
&= -\sigma \ln (P_{t,i} / P_{s,i}) +  \gamma_{t,i} \left( CDD_{t,i} + HDD{t,i} \right) \\
&\qquad + \gamma_{s,i} \left( CDD_{s,i} + HDD{s,i} \right)  + \eta \Delta_{t,s} + u_i
\end{align*}

@@

	\bibitem[ORNL(2004)]{ORNL}
	ORNL (2004). Measurement practices for reliability and power quality. A toolkit of reliability measurement practices. U.S. Department of Energy. https://info.ornl.gov/sites/publications/Files/Pub57467.pdf
